How to improve your relationship with your future self.

### 2011-03-21 workflow.Rnw
Continued editing

All results (numbers, comparisons, tables, figures) should arise
generated from code, not series of mouse clicks or copying and
pasting. If I wanted to re-create the figure you created but including
a new variable or specification, I should be able to do so with just a
few edits to the code rather than knowledge of how you used your
pointing device in your graphical user interface.

Using R, for example, I might specify that the file \Verb+fig1.pdf+
was produced by the following commands in a file called \Verb+fig1.R+.

<<eval=FALSE>>=
thedata<-read.csv("Data/thedata.csv")
pdf('fig1.pdf')
please-plot(outcome by explanatory using thedata)
dev.off()
@

Now, in the future if I wonder how ``that plot on page 10'' was
created, if I suitably commented my manuscript or inserted text into
my MANIFEST.txt file I will know (1) ``that plot'' is from a file
called ``fig1.pdf'' and (2) fig1.pdf was created in
\Verb+figs1and2.R+. In the future, if I wanted to quickly change
fig1.pdf, I could edit the commands quickly and easily to do
so. Another good practice is to name the files descriptively, so that
we don't in fact have ``fig1.pdf'' but
``political-participation-by-education-in-the-south.pdf''. Now, if Figure is
moved to an appendix, the file name is still meaningful.

% An even nicer way to ensure that figures and tables are reproducible
% or easily modifiable by your older and wiser but more harried self is
% to use literate programming. For example, to produce the
% Figure~\ref{fig:giniprot}, I insert the following chunk of code into
% the LaTeX document, and then I asked R to process the document, and
% then produce a pdf file in which the code has disappeared to be
% replaced with an actual figure.\footnote{One can insert the same code
%   chunk in an OpenOffice document and use the odfWeave package to
%   produce OpenOffice documents in which figures and tables replace
%   code. One may also use the LyX interface to LaTeX to write LaTeX
%   documents within a WYSIWYG environment.}

<<fig1code,echo=TRUE,fig=FALSE>>=
##  Make a scatterplot of Protest by Inequality
with(good.df,plot(Gini2004/100,protac2000,
                  xlab='Gini Coefficient 2004 (UNDP)',
                  ylab='Mean Protest Activities\n(World Values Survey 1980-2000)',
                  cex=.8))

##  Label a few interesting points
with(good.df[c("EGY","JOR","USA","SWE","CHL"),],
     text(Gini2004/100,protac2000,labels=Nation,srt=0,cex=.6,pos=3,offset=.1))
@

\begin{figure}[H]
  \begin{center}
<<fig1plot,echo=FALSE,fig=TRUE,width=3,height=3>>=
par(bty="n",xpd=TRUE,pty="s",tcl=-.25)
<<fig1code>>
@
    \caption{Protest activity by income inequality \citep[from][]{norris2009data}.}\label{fig:giniprot}
  \end{center}
\end{figure}

The principle here is to know where tables and figures came from ---
to know what commands created which figures and tables.

\section{No data analyst is an island.}
% MMF: My original title before I saw you had you own comment section
% \section{Comment on your own work. }

Comments, or text inside of a file that is not executed, are a message to your
future self and other consumers of your work. 
Read \citet{nagler1995coding}. Comment your code. Luckily, if you are
using a literate programming practice (Sweave, odfWeave, using R2HTML,
etc..), you can write paragraphs to surround your code as well a
technical comments in the code itself. If you are not using a strictly
literate programming practice then use whatever commenting protocol
exists in your analysis language. Here are some examples:
\begin{verbatim}
# This is a comment in R.
# With a a second line.

@@

A given analysis of a given collection is a long series of
decisions. You need to be able to justify them. Often you justify them
in the body of the paper you are writing. Some decisions will be too small and
technical for inclusion in the paper itself. These need to be
documented in the code itself.

@@

The principle regarding data is to know where the data came from and
what operations were performed on which set of data. I have seen
command files which do not begin with a call to load data, and I have
despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if
I do not know where var1 var2 come from.\footnote{The command
  \Verb+please-fit+ and some other R functions used in this essay come
  from the \Verb+MayIPleaseDoStatistics+ package which emphasizes
  politeness in data analysis. The commands \Verb+please-fit+ and
  \Verb+please-plot+ can be blocked and \Verb+would-you-please-fit+ or
  \Verb+may-I-please-have-a-plot+ can be required using
  \Verb+options(politeness=99)+ }

>>>>

All results (numbers, comparisons, tables, figures) should arise from
code, not series of mouse clicks or copying and pasting. If I wanted
to re-create the figure you created but including a new variable or
specification, I should be able to do so with just a few edits to the
code rather than knowledge of how you used your pointing device in
your graphical user interface.

Using R, for example, I might specify that the file \Verb+fig1.pdf+
was produced by the following commands in a file called
\Verb+fig1.R+.\footnote{The command \Verb+please-plot+ and some other
  R functions used in this essay come from the
  \Verb+MayIPleaseDoStatistics+ package which emphasizes politeness in
  data analysis. The short hand commands like \Verb+please-plot+ can
  be blocked and \Verb+may-I-please-have-a-plot+ can be required using
  \Verb+options(politeness=99)+ }

\begin{Verbatim}[fontsize=\small]
thedata<-read.csv("Data/thedata.csv")
pdf('fig1.pdf')
please-plot(outcome by explanatory using thedata. red lines please.)
dev.off()
\end{Verbatim}

Now, in the future if I wonder how ``that plot on page 10'' was
created: (1) ``that plot'' is from a file called ``fig1.pdf'' and
fig1.pdf was created in \Verb+fig1.R+. In a future where R still
exists, if I wanted to change the figure, I could edit the commands
quickly and easily to do so (since R commands are written in plain
text, and plain text is a format that will be around for a long time
to come).

\paragraph{Principle} Know the provenance of your results so that your
future self or current collaborators can quickly and easily reproduce
your work (and make changes).% --- to know what
% commands created which figures and tables.

\section{No data analyst is an island.}
Data analysis involves a long series of decisions. Each decision
requires justification.  The body of the paper or book or report will
justify the most important decisions. Some decisions will be too small
and technical for inclusion in the paper itself. These need to be
documented in the code itself \citep{nagler1995coding}.

Comments --- unexecuted text inside of a script --- are a message to
your future self and other consumers of your work.  Read
. Comment your code. Luckily, if you are using
a literate programming practice (Sweave, odfWeave, using R2HTML,
etc..), you can write paragraphs to surround your code as well a
technical comments in the code itself. If you are not using a strictly
literate programming practice then use whatever commenting protocol
exists in your analysis language. Here are some examples:
\begin{verbatim}
# This is a comment in R.
# With a a second line.

@@

@@

The principle regarding data is to know where the data came from and
what operations were performed on which set of data. I have seen
command files which do not begin with a call to load data, and I have
despaired. What is the meaning of \Verb+please-fit(var1 with var2)+ if
I do not know where var1 var2 come from.
